### 2013-04-26 docs/mdm_proposal_may2013.tex
added kvj and allyson changes

\documentclass[preprint]{aastex}

\begin{document}
\title{Probing the Milky Way's Dark Matter Halo with RR Lyraes in Tidal Streams I: The Sagittarius Leading Arm}
\author{Adrian Price-Whelan\altaffilmark{1}, Kathryn Johnston\altaffilmark{1}, Allyson Sheffield\altaffilmark{1}}
\altaffiltext{1}{Department of Astronomy, Columbia University, Mail Code 5246, New York, NY 10027}
Galaxies, by mass, are mostly composed of \emph{Dark Matter}. Simulations of structure and galaxy formation -- both with
and without baryons -- lead us to believe that such \emph{halos} of
dark matter have universal density profiles \citep{nfw96} and are triaxial
in shape \citep{jing02}. Including baryons in such simulations tends to
soften the triaxiality and alters the radial density profile. Thus,
precise measurements of the shape, orientation, and radial profile of
dark matter halos provides information about both dark matter physics
and the baryonic processes that helped shape them. Tidal streams from
infalling satellite galaxies have been used to constrain properties of
the Milky Way's halo \citep[e.g.][]{law10}; typically found outwards of 20~kpc, the
orbits of associated stars are still largely under the influence of
the dark matter and thus serve as potentiometers \citep[e.g.][]{johnston99}. Using an updated
method that utilizes full phase-space (6D) kinematic information for a
small sample ($\sim$100s) of stars stripped from a known progenitor
(e.g., Sagittarius; Price-Whelan et al 2013, in prep), it is
possible to accurately constrain the axis ratios, total mass, and
orientation of the Milky Way's dark matter halo. This method relies on a technique for deriving
distances to individual RR Lyrae stars with 2\% accuracy \citep{},
along with anticipated GAIA proper motion errors of
$\sim10-20~\mu\mathrm{as}/\mathrm{yr}$ for such stars at $\sim$
20-30~kpc. This leaves one remaining velocity component -- the radial
velocity.  We plan to observe a
sample of RR Lyrae stars selected to be likley members of Sgr's debris system 
using catalogs of RRab stars from
time-domain, photometric surveys such as LINEAR \citep{}, PTF \citep{}
and CSS \citep{}. 
Our observations will both confirm membership of the debris system in order to construct a target list for Spitzer (proposals are due in August 2013) as well
as measure the missing velocity dimension.

We propose to measure radial velocities of RR Lyraes in the Sgr leading arm using ModSpec on the 2.4m Hiltner
telescope at MDM.  Using the 600 groove mm$^{-1}$ grating blazed at 5000 \AA\ and a 2048x2048 detector with 15 micron pixels (Wilbur), we will have wavelength coverage of about 1000 \AA\, including the H$\beta$ line at 4861 \AA\ and the [OI] sky line at 5577 \AA\ (the latter to aid in the radial velocity determination by identifying any systematic offsets).  We will collect bias and flat field frames at the beginning and end of the night, and ThAr arc lamp spectra for wavelength calibration will be taken throughout the night to account for telescope flexure.  Radial velocity standards will be selected that are of a similar spectral type as the targets (e.g., HD122693  -- a dwarf of spectral type F8). 
Typical equatorial coordinates for the targets are ($\alpha,\delta$)=(13$^{h}-16^{h}$,-5$^{\circ}- +25^{\circ}$); stars in this range will be at airmass $<2$ all night.
The typical $R$-band magnitude is 16.  To achieve S/N of 30 per pixel with the proposed set-up, an integration time of 1200 seconds is needed for a target with $R$=16.  Considering the total amount of observing time of 9 hours per night and the overhead needed for calibration frames, we anticipate observing approximately 40 RR Lyraes.  Reductions will be carried out by Sheffield using a well-developed reduction pipeline for processing ModSpec data.

\begin{thebibliography}[99]
\bibitem[Jing 
\& Suto(2002)]{jing02} Jing, Y.~P., \& Suto, Y.\ 2002, \apj, 574, 538 
\bibitem[Navarro et al.(1996)]{nfw96} Navarro, J.~F., Frenk, 
C.~S., \& White, S.~D.~M.\ 1996, \apj, 462, 563 
\end{thebibliography}

\end{document}